Passenger Prediction in Shared Accounts for Flight Ticket Recommendation

### 2016-06-28 Paper.tex
Add files via upload

\documentclass{llncs}
\bibliographystyle{splncs}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=blue,citecolor=blue]{hyperref}
\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{document}

@@

\maketitle
\begin{abstract}
Personalized recommendation is needed for online flight ticket booking service because it is a difficult task for a traveller to select the flight when the number of available flights is large. Traditionally, we can recommend flights for a user based on his historical orders from his account. However, people often book tickets for companions through his account. In this case, the preferences of other travellers should also be considered. Unfortuanately, before placing the order, people will not provider passengers' information. Therefore, we propose a probabilistic method for passenger prediction based on historical behaviors and contextual knowledge. We then experimentally demonstrate its effectiveness on a real dataset. The result shows that our method outperforms conventional methods.\\
\keywords{recommender system, probabilistic topic model, shared account}
\end{abstract}

\section{Introduction}
\label{sec:intro}
Recent years, with the rapid development of online travel agency, there are more and more passengers booking flight tickets through online service providers. Typically, an user may input his location, destination and departure time, then he will get a list of candidate flights. In general, there are dozens of items within a list, all of them are isomorphic. Thus, it takes quite long time for the user to compare and choose the most appropriate one. Therefore, a recommender system is quite necessary to be applied for better user experience. Figure \ref{fig:res} shows a list for all candidate flights departure from Beijing to Hong Kong when input a query. There are total 156 matchable flights in result. The website only provides simple sorting strategies such as departure time, arrival time, price. Thus, it may take quite long time for users to find the appropriate one.
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.48]{flight_tickets.eps}
\caption{Search result for flight tickets}
\label{fig:res}
\end{figure}\par
Essentially, an item recommender system is used for generating a personalized ranking on a set of items\cite{lv:rec}. Typical recommender systems assume that every account represents a single preference model. However, in online flight ticket booking, it is quite common that a user books tickets for the whole family or companions.  Obviously, every passenger has his own preferences. Unfortuanately, before placing the order, people will not provider passengers' information.  Therefore, if we want to improve the accuracy of recommendation, we have to predict who are going to fly.  \par
 +To predict who is using the account to place orders is regarded as shared account problem. Some approaches have been proposed to extract implicit user information within an account. For example,  in \cite{kabbur:nlmf}\cite{jason:embedding} nonlinear latent factorization is applied to  extract user information. However, these methods only rely on the analysis of relationships between products and users \cite{koen:top-n}\cite{yutaka:modeling}, and context information that can help differentiate users is neglected. Besides, current methods always treat an account as a fixed number of implicit users, and the experiment result can be only evaluated on artificial composite datasets. 
 +The flight booking service providers can acquire user's personal information only after he has submitted the order. Thus, our goal is to predict distribution of passengers in current session based on historical behavior and contextual knowledge so that we can make more pertinent recommendations. To achieve this target, we adopt the probabilistic author topic model to analyze passengers' behaviors. We regard all orders in an account as corpus, all passengers appearing in the account as authors, orders as documents, content attributes of the items and contextual knowledge as item list. Note that if an order has more than one passenger, they are co-authors of this document.\par
 +In experiment section, we apply our method to a real flight ticket order dataset.  And the result shows that our method has higher identification accuracy and a better recommendation accuracy.\par
 +This paper has following contributions:\\
\begin{enumerate}
+\item We provide a probabilistic passenger identification method for online flight booking service.
+\item We apply the passenger prediction results to flight recommendation.
+\item We evaluate our approach on a real flight order dataset, and the result shows that our approach can effectively improve the recommendation accuracy.
\end{enumerate}

\textbf{Paper organization.} The rest of paper is structured as follows. We introduce related work in Section \ref{sec:relw}. In Section \ref{sec:mod}, we define the problem and describe the model. Besides, we also discuss how to do parameter inference. In section \ref{sec:exp}, we introduce the experiment dataset, evaluation metrics and the experimental results of passenge prediction and flight recommendation. Finally, we conclude the paper in section \ref{sec:con}.

\section{Related Work}
\label{sec:relw}
\textbf{Content based recommendation.} In online travel service industry such as airline tickets or hotel booking, the item information is usually highly structured, however, this information is quite mutable. Content-based recommender systems \cite{tech:cset}\cite{lops:handbook}are appropriate for highly structured items. The method needs proper representations of the attributes of items and the profile of user interests, then it tries to match  the profile of user's interests with attributes of items. In general, a content-based recommender system consists of three components. A content analyzer extracts structured attributes of items through a series of pre-processing steps. It produces the inputs for the system, which always need domain related knowledge. A profile learner analyzes and constructs user's profile of interests by collecting user's historical behaviors. In addition, this component also relies on feedbacks of users. A filtering component evaluates the relevance of items to particular users by matching them with users' profiles. The output for an item can be either binary or continuous, in this stage, some similarity measurements including cosine or euclidean distance can usually be applied.\par
\textbf{Flight Ticket Recommendation.} There is little work focusing on personalized flight ticket recommendation. Many online flight ticket booking agencies provide sorting, filtering strategies to improve user experience, such as sorting by price or takeoff time and filtering by airline or departure airport. Others also provide additional information that can help users, for example, placing the hottest flights at top or highlighting some of them. These methods do make sense in helping users find appropriate flights within less time. However, they can not provide personlized recommendations. In 2004, Lorcan Coyle\cite{lor:flight} proposed a Personal Travel Assistant(PTA). PTA is a method of case based reasoning, which takes user's historical orders as cases and recommends the most similar candidate flight to user's query. In addition, PTA also proposes attribute weight training and case storage method.\par
\textbf{Implicit feedback.} The user's feedback can be divided into implict  or explict type. The explicit feedback can be acquired from users' ratings or text comments. However,  users may not willing to provide explict feedback since it needs additional efforts. Moreover, in many applications, a single numeric rating may not be adequate for describing users feeling to an item. Implicit feedback methods\cite{song:mining} try to infer users' attitudes to an item based on his behaviors, such as clicking ,collecting or bookmarking et al. It has the advantage of being free from the problem of lacking feedback data. Nevertheless, implicit feedback based recommender suffers from problems of lack of negative feedbacks and data noises. In order to provide recommendations when only implict feedbacks are available, some methods are proposed, such as one class collaborative filtering and content based recommender.\par

\textbf{Probabilistic topic models.}  Topic models\cite{blei:lda}\cite{mic:atm}\cite{mark:prob} are proposed for automated extraction of useful semantic information from corpus data, which are widely used in information retrieval, document annotation, etc. The main step is extracting latent factors from corpus, named topics, which are commonly probabilistic distribution over words. Topic models have three major advantages over other document modeling methods. At first, topic models provide a completely unsupervised approach to extract topics, thus requiring no document labels and no initialization. Unsupervised methods are necessary for modeling large document corpus, besides, in some fields, with the rapid change of content's topics, predefined topic categories may not reflect the development of specific field. Secondly, latent topics are interpretable respectively, so that extracted results are understandable for users. Finally, each document may consist of multiple topics, document can be considered as probabilistic distribution over topics. Many existing models regard documents as mixtures of topics, such as pLSI, LDA. The author topic model is a generative model that extends LDA to include authorship information for document modeling. In our approach, topic model is innovatively applied to indentify users. \par

@@

\section{User Prediction Model}
\label{sec:mod}
In this Section, we propose a generic model for predicting the probabilistic distribution of passengers under certain contextual environment. Finally we integrate user prediction into recommendation to get higher accuracy. Figure 2 shows an overview of the recommendation process. At first,a dataset for orders are stored, each account may contains several passengers. We use this data to train user prediction model as well as extract preference at user granularity. Before a recommendation process starts, we calculate user distribution probability applying trained model and contextual knowledge of current session. At the final step, we generate a ranked candidate list as the recommendation result based on user prediction result and user's preference.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.40]{overview.eps}
\caption{Overview of flight recommendation based on passenger prediction}
\label{fig:over}
\end{figure}\par

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items, it represents the latent feature. And an account is a multinomial distribution over topics, it can represent the preference of an account. Each purchase can be regared a sample that selects a topic z from account a, and takes an item from topic z.\par
When topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite different to be modeled. Conventional items only have static features, means that the content of an item is less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, so we can't consider the ticket of different price as one item. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly.\par
Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. Some factors are discrete while others are continuous. For continuous factors, we can partition them into discrete non-overlapping intervals applying . Thus, the number of discrete alternatives of each factor is also very limited. We take orders within each account as a corpus. Our notations are summarized in table \ref{tab:not}.\par
\begin{table}[!htbp]
\centering
\caption{\label{tab:not}Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected Factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

@@

\section{Experiments}
\label{sec:exp}
In the following subsections, we introduce datasets, preference-based recommendation method and evaluation metrics. Then we evaluate the performance of user prediction and passenger prediction based recommendation respectively.
\subsection{Dataset and preference-based recommendation method}
In our experiments, we use a real flight ticket dataset, consisting of all submitted orders for two years. An order contains account id and all passengers' id. We use desensitized passenger id to distinguish individuals from each other. Besides, an order contains some travel information that can be considered as decisive factors, such as departure airport, arrival airport, order submitted time, takeoff time, etc. In our experiment, we select active users whose amount of history orders reaches a threshold, thus user's preference can be extracted with more confidence. In addition, many researches\cite{amy:guess}\cite{yutaka:modeling} on shared account recommendation generate an artificial dataset by compossing single accounts' data. We also generate an artificial dataset combining single passengers from different accounts. We extract two passengers from arbitrary accounts and combine them as one account. Overviews of two datasets and the meta data of orders are summarized in table \ref{tab:dat} and table \ref{tab:mdat}.
\begin{table}[!htbp]
\centering
\caption{\label{tab:dat}Datasets}
\begin{tabular}{|c|c|c|c|} \hline
& \# accounts & \# passengers & \# orders\\ \hline
Real & 4632&7034&38907 \\ \hline
Artificial &1604&3208&29759\\ \hline
\end{tabular}
\end{table}

@@

In addition, different users may cares different factors, thus we introduce a weight vector for every account. The evaluation of this weight vector contains two steps, initializing and training, respectively. At the initializing step, we estimate how much a user focuses on a factor through the information entropy, as mentioned in \cite{yang:pers}. We believe that if a user focuses on a factor, his behavior will be less random. The entropy $H$ is a measure to describe the uncertainty for discrete random variable $X$ and probability mass function $P(X)$. Here is the mathematical representation for entropy.\\
\begin{equation}
H(X)=E[-lnP(X)]= - \sum_{i-1}^n P(x_i)log_bP(X_i)
\end{equation}\par
The value $b$ is the base of logarithm, and is usually 2. The entropy of every factor can be calculated and initialized as the weight value after normalization. Considering that different factors have a variety quantity of alternatives and the user may concern different factors some time later, we update the weight vector once the user submits an order. For each submission, we have already produced a ranked list for recommendations, if the user doesn't choice an item (i.e, a ticket) from top-$N$ items, we consider this recommendation fails, so that the weight vector need be retrained to make the rank of selected item improved. The concept of weight training is mentioned in \cite{lor:flight}. By performing weight training for factors, we hope that the result of recommendation getting better in the future. The training procedure is quite straightforward. At first a learning rate should be set, such as $\alpha = 0.1$ . If the selected item has lower similarity for factor $f$ than items ranked before it, the weight value of this factor should be decreased. Otherwise, the weight value should be increased. To avoid the problem of over-fitting, we define a maximum number of iteration step. So the stop criteria is either $1)$ the rank of selected items can't be improved anymore or $2)$ the procedure reaches the iteration time limitation.\\
 +The preference based recommendation algorithm is described as Algorithm \ref{alg:3}. First extracting a user's preference  $P$ and initializing weighted list $W$ through entropy. Items in candidate list are ranked based on inner product of $W$ and corresponding value of factor. After every recommendation, the list is retrained and stored depending on the item that user actually chosen.

@@

\subsection{Settings and Evaluation Metrics}
Essentially, flight ticket recommendation is an approach to select the best mached ticket by measuring the similarities between candidate tickets and user's preferences. We take the latest order of a user as the test data and the rest ones as the training data. The test data also plays a role of providing contextual knowledge for user prediction.\par
One significant issue is that we can't know real candidates at the time when user performed the search action. Fortunately, we can get an approximate set of candidate by collecting orders from all users in dataset with the same order date, takeoff date, departure city and arrival city as the test order, the granularity of the simulate candidate is to every flight, class and price range. Price range is a discrete interval representing the degree of the price based on domain knowledge. In order to get a persuasive evaluation result, we filter out test orders with the size of candidate list less than 20, and the average size of candidates is 45.\par
We apply two commonly used metrics MAP(Mean Average Precision) and top-$N$ hit rate to evaluate the recommendation performance. The definition of MAP is described as follow:\\
\begin{equation}
MAP = \frac{\sum_{i=1}^{|M|}Acc(u_i)}{|M|}
\end{equation}

@@

where $|M|$ is the number of total test orders, $Acc(u_i)$ represents the accuracy of recommendation based on the rank percentages of test orders. The metric MAP is the average of each user's accuracy, and is a measurement for general performance. The top-$N$ hit rate is the actual strategy for recommendation, we can stick up top-$N$ candidates based on the ranked list. For each purchase, if top-$N$ candidates hit the chosen order, we consider this recommendation has accuracy 1, else the accuracy is 0. We also calculate MAP for all accounts \par
For passenger prediction, we propose the following metric to evaluate the accuracy:\\
\begin{equation}
\label{eq:10}
Acc(u_i) = \frac{P' \cap A}{\sqrt{|P'| \times |A|}}
\end{equation}\par
Since we treat every account as a corpus, so a probabilistic distribution list is generated for all passengers contained in the account. At the first stage, we need to decide the passenger number for each test order. Assuming that there are total $|P|$ passengers in an account. So in average case, the probability of every passenger is $P_A = \frac{1}{|P|}$. We decide the passenger number by omitting passengers whose probability is not greater than $P_A$. We denote predicted passenger list as $P'$, actual passenger list as $A$. Thus we can get the prediction accuracy by Equation \ref{eq:10}. The numerator is the number of hitted passengers, the denominator is penalty item for the size of predicted passenger list and predicted passenger list.

\subsection{Passenger Prediction}
We train a probabilistic model for every account where topic number selected varying ${5,10,\dots,50}$. We pick a test order for each account as a new document, then predict every passenger's probability of placing this order by Equation \ref{eq:5}. According to above evaluating metrics, the mean prediction accuracy rate of (a) real dataset and (b) artificial dataset is shown in Figure \ref{fig:pred}. The left sub-figure is the evaluation result on the real dataset and the right one is on the artificial dataset. The figure's horizontal axis shows the number of trained topics and the vertical axis shows the prediction accuracy computed by applying above mentioned metrics. The baseline means assigning random probabilities to passengers, we also record $10$ batch of random results as our approach does. For the artificial dataset, there are two passengers in every account and one passenger for every test order, so the accuracy of prediction is either 1.0 or 0. The results demonstrate that our model has a higher prediction precision than the baseline method on both datasets. We notice that the average improvements of prediction accuracy on artificial dataset is higher than on the real dataset. Besides, when the topic number is 10 or 15, both datasets get the best performance while the performance goes down with the topic number increasing. The sizes of corpus' vocabulary and number of words in each order also impact the optimal topic number. In the next subsection, we evaluate the recommendation performance with topic number fixed at 10.\par

\begin{figure}[!h]
              \begin{minipage}[t]{0.47\linewidth}
              \centering
              \includegraphics[width=1.17\linewidth]{e5.eps}\\
              \end{minipage}
              \begin{minipage}[t]{0.005\linewidth}~~~
              \end{minipage}
              \begin{minipage}[t]{0.47\linewidth}
              \centering
              \includegraphics[width=1.17\linewidth]{e6.eps}\\
              \end{minipage}
              \begin{minipage}[t]{0.005\linewidth}~~~
              \end{minipage}
              \caption{MAP of user prediction}
          \label{fig:pred}
\end{figure}

>>>>

\documentclass{llncs}
\bibliographystyle{splncs}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=blue,citecolor=blue]{hyperref}
\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{document}

@@

\maketitle
\begin{abstract}
Personalized recommendation is needed for online flight ticket booking service because it is a difficult task for a traveller to select the flight when the number of available flights is large. Traditionally, we can recommend flights for a user based on his historical orders from his account. However, people often book tickets for companions through his account. In this case, the preferences of other travellers should also be considered. Unfortunately, before placing the order, people will not provider passengers' information. Therefore, we propose a probabilistic method for passenger prediction based on historical behaviors and contextual knowledge. We then experimentally demonstrate its effectiveness on a real dataset. The result shows that our method outperforms conventional methods.\\
\keywords{recommender system, probabilistic topic model, shared account}
\end{abstract}

\section{Introduction}
\label{sec:intro}
Recent years, with the rapid development of online travel agency, there are more and more passengers booking flight tickets through online service providers. Typically, an user may input his location, destination and departure time, then he will get a list of candidate flights. In general, there are dozens of items within a list, all of them are isomorphic. Thus, it takes quite long time for the user to compare and choose the most appropriate one. Therefore, a recommender system is quite necessary to be applied for better user experience. Figure \ref{fig:res} shows a list for all candidate flights departure from Beijing to Hong Kong when input a query. There are total 156 matchable flights in result. The website only provides simple sorting strategies such as departure time, arrival time, price. Thus, it may take quite long time for users to find the appropriate one.
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.48]{flight_tickets.eps}
\caption{Search result for flight tickets}
\label{fig:res}
\end{figure}\par
Essentially, an item recommender system is used for generating a personalized ranking on a set of items\cite{lv:rec}. Typical recommender systems assume that every account represents a single preference model. However, in online flight ticket booking, it is quite common that a user books tickets for the whole family or companions.  Obviously, every passenger has his own preferences. Unfortunately, before placing the order, people will not provider passengers' information.  Therefore, if we want to improve the accuracy of recommendation, we have to predict who are going to fly.  \par
 To predict who is using the account to place orders is regarded as shared account problem. Some approaches have been proposed to extract implicit user information within an account. For example,  in \cite{kabbur:nlmf}\cite{jason:embedding} nonlinear latent factorization is applied to  extract user information. However, these methods only rely on the analysis of relationships between products and users \cite{koen:top-n}\cite{yutaka:modeling}, and context information that can help differentiate users is neglected. Besides, current methods always treat an account as a fixed number of implicit users, and the experiment result can be only evaluated on artificial composite datasets. 
 The flight booking service providers can acquire user's personal information only after he has submitted the order. Thus, our goal is to predict distribution of passengers in current session based on historical behavior and contextual knowledge so that we can make more pertinent recommendations. To achieve this target, we adopt the probabilistic author topic model to analyze passengers' behaviors. We regard all orders in an account as corpus, all passengers appearing in the account as authors, orders as documents, content attributes of the items and contextual knowledge as item list. Note that if an order has more than one passenger, they are co-authors of this document.\par
 In experiment section, we apply our method to a real flight ticket order dataset.  And the result shows that our method has higher identification accuracy and a better recommendation accuracy.\par
 This paper has following contributions:\\
\begin{enumerate}
\item We provide a probabilistic passenger identification method for online flight booking service.
\item We apply the passenger prediction results to flight recommendation.
\item We evaluate our approach on a real flight order dataset, and the result shows that our approach can effectively improve the recommendation accuracy.
\end{enumerate}

\textbf{Paper organization.} The rest of paper is structured as follows. We introduce related work in Section \ref{sec:relw}. In Section \ref{sec:mod}, we define the problem and describe the model. Besides, we also discuss how to do parameter inference. In section \ref{sec:exp}, we introduce the experiment dataset, evaluation metrics and the experimental results of passenger prediction and flight recommendation. Finally, we conclude the paper in section \ref{sec:con}.

\section{Related Work}\par
\label{sec:relw}
\textbf{Content based recommendation.} In online travel service industry such as airline tickets or hotel booking, the item information is usually highly structured, however, this information is quite mutable. Content-based recommender systems \cite{tech:cset}\cite{lops:handbook}are appropriate for highly structured items. The method needs proper representations of the attributes of items and the profile of user interests, then it tries to match  the profile of user's interests with attributes of items. In general, a content-based recommender system consists of three components. A content analyzer extracts structured attributes of items through a series of pre-processing steps. It produces the inputs for the system, which always need domain related knowledge. A profile learner analyzes and constructs user's profile of interests by collecting user's historical behaviors. In addition, this component also relies on feedbacks of users. A filtering component evaluates the relevance of items to particular users by matching them with users' profiles. The output for an item can be either binary or continuous, in this stage, some similarity measurements including cosine or euclidean distance can usually be applied.\par
\textbf{Flight Ticket Recommendation.} There is little work focusing on personalized flight ticket recommendation. Many online flight ticket booking agencies provide sorting, filtering strategies to improve user experience, such as sorting by price or takeoff time and filtering by airline or departure airport. Others also provide additional information that can help users, for example, placing the hottest flights at top or highlighting some of them. These methods do make sense in helping users find appropriate flights within less time. However, they can not provide personalized recommendations. In 2004, Lorcan Coyle\cite{lor:flight} proposed a Personal Travel Assistant(PTA). PTA is a method of case based reasoning, which takes user's historical orders as cases and recommends the most similar candidate flight to user's query. In addition, PTA also proposes attribute weight training and case storage method.\par
\textbf{Implicit feedback.} The user's feedback can be divided into implicit  or explicit type. The explicit feedback can be acquired from users' ratings or text comments. However,  users may not willing to provide explicit feedback since it needs additional efforts. Moreover, in many applications, a single numeric rating may not be adequate for describing users feeling to an item. Implicit feedback methods\cite{song:mining} try to infer users' attitudes to an item based on his behaviors, such as clicking ,collecting or bookmarking et al. It has the advantage of being free from the problem of lacking feedback data. Nevertheless, implicit feedback based recommender suffers from problems of lack of negative feedbacks and data noises. In order to provide recommendations when only implicit feedbacks are available, some methods are proposed, such as one class collaborative filtering and content based recommender.\par

\textbf{Probabilistic topic models.}  Topic models\cite{blei:lda}\cite{mic:atm}\cite{mark:prob} are proposed for automated extraction of useful semantic information from corpus data, which are widely used in information retrieval, document annotation, etc. The main step is extracting latent factors from corpus, named topics, which are commonly probabilistic distribution over words. Topic models have three major advantages over other document modeling methods. At first, topic models provide a completely unsupervised approach to extract topics, thus requiring no document labels and no initialization. Unsupervised methods are necessary for modeling large document corpus, besides, in some fields, with the rapid change of content's topics, predefined topic categories may not reflect the development of specific field. Secondly, latent topics are interpretable respectively, so that extracted results are understandable for users. Finally, each document may consist of multiple topics, document can be considered as probabilistic distribution over topics. Many existing models regard documents as mixtures of topics, such as pLSI, LDA. The author topic model is a generative model that extends LDA to include authorship information for document modeling. In our approach, topic model is innovatively applied to identify users. \par

@@

\section{Passenger Prediction Model}
\label{sec:mod}
In this Section, we propose a generic model for predicting the probabilistic distribution of passengers under certain contextual environment. Finally we integrate user prediction into recommendation to get higher accuracy. Figure 2 shows an overview of the recommendation process. At first,a dataset for orders are stored, each account may contains several passengers. We use this data to train user prediction model as well as extract preference at user granularity. Before a recommendation process starts, we calculate user distribution probability applying trained model and contextual knowledge of current session. At the final step, we generate a ranked candidate list as the recommendation result based on user prediction result and user's preference.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.40]{overview.eps}
\caption{Overview of flight recommendation based on passenger prediction}
\label{fig:over}
\end{figure}\par

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items, it represents the latent feature. And an account is a multinomial distribution over topics, it can represent the preference of an account. Each purchase can be regarded a sample that selects a topic z from account a, and takes an item from topic z.\par
When topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite different to be modeled. Conventional items only have static features, means that the content of an item is less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, so we can't consider the ticket of different price as one item. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly.\par
Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. Some factors are discrete while others are continuous. For continuous factors, we can partition them into discrete non-overlapping intervals applying . Thus, the number of discrete alternatives of each factor is also very limited. We take orders within each account as a corpus. Our notations are summarized in table \ref{tab:not}.\par
\begin{table}[!htbp]
\centering
\caption{\label{tab:not}Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected Factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

@@

\section{Experiments}
\label{sec:exp}
In the following subsections, we introduce datasets, preference-based recommendation method and evaluation metrics. Then we evaluate the performance of user prediction and passenger prediction based recommendation respectively.
\subsection{Dataset and preference-based recommendation method}
In our experiments, we use a real flight ticket dataset, consisting of all submitted orders for two years. An order contains account id and all passengers' id. We use desensitized passenger id to distinguish individuals from each other. Besides, an order contains some travel information that can be considered as decisive factors, such as departure airport, arrival airport, order submitted time, takeoff time, etc. In our experiment, we select active users whose amount of history orders reaches a threshold, thus user's preference can be extracted with more confidence. In addition, many researches\cite{amy:guess}\cite{yutaka:modeling} on shared account recommendation generate an artificial dataset by composing single accounts' data. We also generate an artificial dataset combining single passengers from different accounts. We extract two passengers from arbitrary accounts and combine them as one account. Overviews of two datasets and the meta data of orders are summarized in table \ref{tab:dat} and table \ref{tab:mdat}.
\begin{table}[!htbp]
\centering
\caption{\label{tab:dat}Datasets}
\begin{tabular}{|c|c|c|c|} \hline
& \# accounts & \# passengers & \# orders\\ \hline
Real & 4632&7034&38907 \\ \hline
Artificial &1604&3208&29759\\ \hline
\end{tabular}
\end{table}

@@

In addition, different users may cares different factors, thus we introduce a weight vector for every account. The evaluation of this weight vector contains two steps, initializing and training, respectively. At the initializing step, we estimate how much a user focuses on a factor through the information entropy, as mentioned in \cite{yang:pers}. We believe that if a user focuses on a factor, his behavior will be less random. The entropy $H$ is a measure to describe the uncertainty for discrete random variable $X$ and probability mass function $P(X)$. Here is the mathematical representation for entropy.\\
\begin{equation}
H(X)=E[-lnP(X)]= - \sum_{i-1}^n P(x_i)log_bP(X_i)
\end{equation}\par
The value $b$ is the base of logarithm, and is usually 2. The entropy of every factor can be calculated and initialized as the weight value after normalization. Considering that different factors have a variety quantity of alternatives and the user may concern different factors some time later, we update the weight vector once the user submits an order. For each submission, we have already produced a ranked list for recommendations, if the user doesn't choice an item (i.e, a ticket) from top-$N$ items, we consider this recommendation fails, so that the weight vector need be retrained to make the rank of selected item improved. The concept of weight training is mentioned in \cite{lor:flight}. By performing weight training for factors, we hope that the result of recommendation getting better in the future. The training procedure is quite straightforward. At first a learning rate should be set, such as $\alpha = 0.1$ . If the selected item has lower similarity for factor $f$ than items ranked before it, the weight value of this factor should be decreased. Otherwise, the weight value should be increased. To avoid the problem of over-fitting, we define a maximum number of iteration step. So the stop criteria is either $1)$ the rank of selected items can't be improved anymore or $2)$ the procedure reaches the iteration time limitation.\\
The preference based recommendation algorithm is described as Algorithm \ref{alg:3}. First extracting a user's preference  $P$ and initializing weighted list $W$ through entropy. Items in candidate list are ranked based on inner product of $W$ and corresponding value of factor. After every recommendation, the list is retrained and stored depending on the item that user actually chosen.

@@

\subsection{Settings and Evaluation Metrics}
Essentially, flight ticket recommendation is an approach to select the best matched ticket by measuring the similarities between candidate tickets and user's preferences. We take the latest order of a user as the test data and the rest ones as the training data. The test data also plays a role of providing contextual knowledge for user prediction.\par
One significant issue is that we can't know real candidates at the time when user performed the search action. Fortunately, we can get an approximate set of candidate by collecting orders from all users in dataset with the same order date, takeoff date, departure city and arrival city as the test order, the granularity of the simulate candidate is to every flight, class and price range. Price range is a discrete interval representing the degree of the price based on domain knowledge. In order to get a persuasive evaluation result, we filter out test orders with the size of candidate list less than 20, and the average size of candidates is 45.\par
We apply two commonly used metrics MAP(Mean Average Precision) and top-$N$ hit rate to evaluate the recommendation performance. The definition of MAP is described as follow:\\
\begin{equation}
MAP = \frac{\sum_{i=1}^{|M|}Acc(u_i)}{|M|}
\end{equation}

@@

where $|M|$ is the number of total test orders, $Acc(u_i)$ represents the accuracy of recommendation based on the rank percentages of test orders. The metric MAP is the average of each user's accuracy, and is a measurement for general performance. The top-$N$ hit rate is the actual strategy for recommendation, we can stick up top-$N$ candidates based on the ranked list. For each purchase, if top-$N$ candidates hit the chosen order, we consider this recommendation has accuracy 1, else the accuracy is 0. We also calculate MAP for all accounts \par
For passenger prediction, we propose the following metric to evaluate the accuracy:\\
\begin{equation}
\label{eq:10}
Acc(u_i) = \frac{P' \cap A}{\sqrt{|P'| \times |A|}}
\end{equation}\par
Since we treat every account as a corpus, so a probabilistic distribution list is generated for all passengers contained in the account. At the first stage, we need to decide the passenger number for each test order. Assuming that there are total $|P|$ passengers in an account. So in average case, the probability of every passenger is $P_A = \frac{1}{|P|}$. We decide the passenger number by omitting passengers whose probability is not greater than $P_A$. We denote predicted passenger list as $P'$, actual passenger list as $A$. Thus we can get the prediction accuracy by Equation \ref{eq:10}. The numerator is the number of hit passengers, the denominator is penalty item for the size of predicted passenger list and predicted passenger list.

\subsection{Passenger Prediction}
We train a probabilistic model for every account where topic number selected varying ${5,10,\dots,50}$. We pick a test order for each account as a new document, then predict every passenger's probability of placing this order by Equation \ref{eq:5}. According to above evaluating metrics, the mean prediction accuracy rate of (a) real dataset and (b) artificial dataset is shown in Figure \ref{fig:pred}. The left sub-figure is the evaluation result on the real dataset and the right one is on the artificial dataset. The figure's horizontal axis shows the number of trained topics and the vertical axis shows the prediction accuracy computed by applying above mentioned metrics. The Random method means assigning random probabilities to passengers, we also record $10$ batch of random results as our approach does. For the artificial dataset, there are two passengers in every account and one passenger for every test order, so the accuracy of prediction is either 1.0 or 0. The results demonstrate that our model has a higher prediction precision than the random method on both datasets. We notice that the average improvements of prediction accuracy on artificial dataset is higher than on the real dataset. Besides, when the topic number is 10 or 15, both datasets get the best performance while the performance goes down with the topic number increasing. The sizes of corpus' vocabulary and number of words in each order also impact the optimal topic number. In the next subsection, we evaluate the recommendation performance with topic number fixed at 10.\par

\begin{figure}[!h]
              \begin{minipage}[t]{0.47\linewidth}
              \centering
              \includegraphics[width=1.145\linewidth]{e5.eps}\\
              \end{minipage}
              \begin{minipage}[t]{0.005\linewidth}~~~
              \end{minipage}
              \begin{minipage}[t]{0.47\linewidth}
              \centering
              \includegraphics[width=1.145\linewidth]{e6.eps}\\
              \end{minipage}
              \begin{minipage}[t]{0.005\linewidth}~~~
              \end{minipage}
              \caption{MAP of user prediction}
          \label{fig:pred}
\end{figure}

\subsection{Recommendation}
After obtaining the predicted probabilistic distribution of passengers, we can extend preference based recommendation approach by only extracting preferences of predicted passengers. If there are multiple predicted passengers for one test order, we generate a composite preference model that combines all involved passenger's preferences together. The probability of every passenger is normalized.\par
We use MAP and top-$N$ hit rate mentioned above to evaluate recommendation performance. We compare different recommendation approaches including preference based recommendation(shorted as Pre), transfered preference based recommendation(tfPre), price rank(pRank), hot rank(hRank) and predicting preference based recommendation(pdPre). The transfered preference based method is the same as the previous one except it takes the difference between flight distribution of various air routes into consideration. This method groups training data by route, and evaluates similarity between routes by flight distribution in some attributes such as airline, class, price range. If a user isn't active in target route, we transfer his preferences from most similar active routes with a transfer rate. In our experiment, the transfer rate is  $ \alpha = 0.5$. Price rank is a simple strategy that ranks candidates by price in an ascending order. Hot rank is a strategy that ranks candidates according to the hotness, which is based on the total order amount within a period of the previous two weeks, items are distinguished by flight number, class and price range.\par
